---
title: "Modeling Team VES Data"
format: html
editor: source
---

## Load Packages

```{r}

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(tidyverse, here, DBI, RPostgres, lubridate)

```

## Connect to `ribbitr` database

```{r}

tryCatch({
    drv <- dbDriver("Postgres")
    print("Connecting to Database…")
    connection <- dbConnect(drv,
                 dbname = Sys.getenv("aws_dbname"),
                 host = Sys.getenv("aws_host"),
                 port = Sys.getenv("aws_port"),
                 user = Sys.getenv("aws_user"),
                 password = Sys.getenv("aws_password"),
                 timezone=NULL)
    print("Database Connected!")
    },
    error=function(cond) {
            print("Unable to connect to Database.")
    })

#search path
dbExecute(connection, "set search_path to survey_data")

```

## Query VES 'flavored' data

### All VES Data

```{r}

ves_q <- "select l.location, r.region, s.site, v.date, s2.detection_type, v2.species_ves, v2.count
          from location l
          join region r on l.location_id = r.location_id
          join site s on r.region_id = s.region_id
          join visit v on s.site_id = v.site_id
          join survey s2 on v.visit_id = s2.visit_id
          join ves v2 on s2.survey_id = v2.survey_id;"

#### tally count per species on grouping of location, region, site, and date
#### add year column for summarization
all_ves_data <- dbGetQuery(connection, ves_q) %>% 
  group_by(location, region, site, date, species_ves) %>% 
  reframe(count = n()) %>% 
  rename(species = species_ves) %>% 
  mutate(year = year(date))

```

### Panama Capture VES Flavored Data

```{r}

p_capture <- "select l.location, r.region, s.site, v.date, v.survey_time, s2.detection_type, v2.species_capture
             from location l
             join region r on l.location_id = r.location_id 
             join site s on r.region_id = s.region_id 
             join visit v on s.site_id = v.site_id 
             join survey s2 on v.visit_id = s2.visit_id 
             join capture v2 on s2.survey_id = v2.survey_id
             where l.location = 'panama';"

#### tally count per species on grouping of location, region, site, and date
#### add year column for summarization
panama_capture_data <- dbGetQuery(connection, p_capture) %>% 
  group_by(location, region, site, date, species_capture) %>% 
  reframe(count = n()) %>% 
  arrange(date) %>% 
  rename(species = species_capture) %>% 
  mutate(year = year(date))


```

#### `rbind` Panama Captures with all VES

```{r}

all_ves_flavored <- rbind(all_ves_data, panama_capture_data)
  

```

## Visualize

### Counts through time - Bar Plot

```{r}

year_reg_data <- all_ves_flavored %>% 
  select(region, year, species) %>% 
  group_by(region, year) %>%
  reframe(count = n()) %>% 
  group_by(year) %>% 
  mutate(total = sum(count))


# Create a custom color palette with 16 unique colors
custom_palette <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00",
                    "#FFFF33", "#A65628", "#F781BF", "#999999", "#66C2A5",
                    "#FC8D62", "#8DA0CB", "#E78AC3", "#A6D854", "#FFD92F", "#B3B3B3")



# Create the ggplot
ggplot(data = year_reg_data, aes(x = factor(year), y = count, fill = region)) +
  geom_col() +

  # Customize the appearance
  theme_minimal() +  # Choose a theme
  labs(
    title = "VES by Year and Region",
    x = "Year",
    y = "Amphibian Visual Encounters",
    fill = "Region"
  ) +
  scale_fill_manual(values = custom_palette) +  # Use custom colors
  theme(
    legend.position = "top",  # Move the legend to the top
    legend.key.size = unit(0.5, "lines"),  # Adjust the size of the legend key
    legend.text = element_text(size = 8),  # Adjust the size of the legend text
    panel.grid.major = element_blank(),  # Remove gridlines
    panel.border = element_blank(),  # Remove panel border
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
  geom_text(aes(x = factor(year), y = total, label = total),
            vjust = -0.5, size = 2, color = "black")

```

## Sierra Nevada

-   Per Site: 1 VES survey + 3 CMR surveys. Usually the day before CMR or the first day of CMR conducted concurrently with CMR.
-   Attempt bi-annual survey period
    -   Summer only
    - each CMR site (of which there are ~50 site IDs located in ~20 basins) is visited twice a summer with visits spaced about a          month apart. The goal is to conduct a 3-day CMR, and to conduct a VES on day 0 (prior to CMR) or day 1 of the primary period        for each visit. The only exception to this study design is Dog lake, which we visit 3-6 times in a summer conducting 1-day CMR       while conducting VES concurrently.

There are a number of sites that we visit annually and a large group of sites (500-800?) that we visit every decade or so where we just conduct VES. 

## Panama

-   Per Site: 3 day surveys + 1 night survey (attempt at every site)
    - capture every organism they see (capture). If not able to capture, record visual observation (VES). Attempt to ID and record every organism heard (aural).
    - capture and VES can be used together as no one single organism would be captured and VES data point.
    - See above query for utilizing both panama capture and VES to get raw counts.
-   Bi-annual survey period (June/July and November/December)
    -   straddles transition period of rainy to dry or dry to rainy season.
    
## Penn

- Per Site: 3 Aural survey + 3 VES surveys + 1 (or 2) Capture surveys
  - 
- Attempt bi-annual survey period
  - Summer to early fall
